---
title: "sensory_hair_analysis"
author: "Andrea Rummel"
date: '2022-10-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '..') 
```

\newpage

Packages required:
```{r part1pkg, results='hide', message=FALSE}
library(ggplot2)
library(caper)
library(ape)
library(phytools)
library(tidyverse)
```

# raw data processing

``` {r}
allregions <- read.csv("melissa_data_quant+initial_qual_22March22.csv")
allregions <- allregions[,c(1, 3:5, 16, 22:30)] 

allregions$Species <- str_replace(allregions$Species," ", "_")
allregions$Species[19:23] <- "Lasiurus_cinereus"
allregions$Species[29:30] <- "Micronycteris_schmidtorum"
allregions$Species[36:40] <- "Myotis_keaysi"
allregions$Species[41:45] <- "Natalus_tumidirostris"
allregions$Species[53:57] <- "Rhinopoma_hardwickii"
allregions$Species[24:28] <- "Megaderma_spasma" # this is to replace lavia frons
allregions$Species[51:52] <- "Pteronotus_davyi"

```

# PCA

# scale data - and center? for PCA
``` {r}
all.scaled <- allregions
for (i in 1:length(colnames(all.scaled))){
  if (is.numeric(all.scaled[,i])==TRUE)
    all.scaled[,i] <- as.numeric(scale(all.scaled[,i], center = TRUE, scale = TRUE))
  else
    all.scaled[,i] <- all.scaled[,i]
}

all.scaled[is.na(all.scaled)] = 0

```

# make phylogenetic tree 
``` {r}
taxa <- read.csv("Focal_Species.csv", row.names =1)
batTree <- read.tree("ShiRabosky.tree")
batSpecies <- rownames(taxa) # makes a vector of your species
batTree <- drop.tip(batTree, batTree$tip.label
                    [-na.omit(match(batSpecies, batTree$tip.label))]) #trims tree to your taxa
mar=c(0.1, 1, 0.1, 0.1)
plotTree(batTree,fsize=0.8, ftype="i", lwd=.9)

```

# make means datafrmae
``` {r}
# make means dataframe
means.scaled <- aggregate(all.scaled[,6:14],list(Species=all.scaled$Species),mean, na.rm=TRUE, na.action = pass)
pancakes <- all.scaled[,c(1:4)]
pancakes <- distinct(pancakes)
sd.all <- aggregate(all.scaled[,6:14], list(Species=all.scaled$Species),sd, na.rm=TRUE)
# add in categorical variables
means.scaled <- cbind(pancakes, means.scaled[, -1])
rownames(means.scaled) <- means.scaled$Species
rownames(means.scaled)

remove(pancakes)

```


# order data so it matches tip labels and make comparative data object
``` {r}
order.species <- batTree$tip.label
df.test <- data.frame(Species = order.species)

means.scaled.sorted <- means.scaled[match(df.test$Species,means.scaled$Species),]
# meansnorm.sorted <- meansnorm.sorted[-7,] # removes desmodus from mean dataframe
meansort.scaled.long <- pivot_longer(means.scaled.sorted, c(5:(ncol(means.scaled.sorted))), 
                                  names_to = "region", values_to = "density")

comp.scaled <- comparative.data(data = means.scaled.sorted, phy = batTree, 
                          names.col = "Species", na.omit = FALSE, warn.dropped = TRUE)

```

# phylogenetic signal
``` {r}
roi <- c("Elbow_cm2","Ankleknee.1", "Propat_cm2", "CD_Plagio_cm2", "MD_Plagio_cm2", "CD_Dactylo_Major_cm2", 
         "MD_Dactylo_Major_cm2", "CD_Dactylo_Medius_cm2", "MD_Dactylo_Medius_cm2")

lambdaresult <- data.frame(roi, 
                           lambda = rep(NA,length(roi)),
                           logL = rep(NA,length(roi)), 
                           logL0 = rep(NA,length(roi)), 
                           lamb_P = rep(NA,length(roi)), 
                           K = rep(NA, length(roi)), 
                           K_P = rep(NA, length(roi)))

# when lambda = 1, high phylogenetic signal. when lambda = 0, branch structure is star phylogeny 
# all regions 
for (i in 5:13) {
  sub.lamb <- phylosig(batTree, means.scaled.sorted[[i]],method = "lambda",test=TRUE)
  sub.k <- phylosig(batTree,means.scaled.sorted[[i]], method = "K", test = TRUE)
  for (j in 1:4) {
    lambdaresult[[i-4,j+1]] <- sub.lamb[[j]]
  }
  for (l in 1:2) {
    lambdaresult[[i-4,l+5]] <- sub.k[[l]]
  }
}

```
only 3 regions are significant

# phylogenetic PCA

```{r}
# phylogenetic pca, method lambda
# here only using 5:13 because not doing proxdist
# temporarily make tadarida propat mean = CD_Plagio mean - need to check on this blank ??
ppca.lambda.scaled <- phyl.pca(batTree, means.scaled.sorted[,c(5:13)], method = "lambda", mode="cov", opt = "ML")
summary(ppca.lambda.scaled) # pc1 = 70% of variance 

# extract pc scores for first two component and add to dat dataframe
means.scaled.sorted$pc1 <- ppca.lambda.scaled$S[,1] # indexing the first column
means.scaled.sorted$pc2 <- ppca.lambda.scaled$S[,2] # indexing the second column

```

# models using PC1 for ecological comparison

all non-significant
```{r}
mod.pca <- pgls(ppca.lambda.scaled$S[, 1] ~ PreyType, comp.scaled)
summary(mod.pca)
anova(mod.pca) # significant result in standard analysis is not significant with phylogenetic analysis

# more closely related species go after similar prey types/modes (this is mobile vs immobile)

comp.scaled %>% pgls(ppca.lambda.scaled$S[, 1] ~ PreyType, data = .) %>% summary
comp.scaled %>% pgls(ppca.lambda.scaled$S[, 1] ~ Guild, data = .) %>% summary


```

# pairwise comparisons among regions with unscaled data - is this the right call?

setup data
```{r}
allregions$prox <- rowMeans(allregions[c(6:7,9:10)])
allregions$dist <- rowMeans(allregions[c(11:14)])
allregions$cran <- rowMeans(allregions[c(6, 10, 12, 14)])
allregions$caud <- rowMeans(allregions[c(7, 9, 11, 13)])

allregions.long <- allregions %>% pivot_longer(cols = c(6:18), values_to = "density", names_to = "region")

```

make means dataframe
```{r}
means.all <- aggregate(allregions[,6:18],list(Species=allregions$Species),mean, na.rm=TRUE, na.action = pass)
pancakes <- allregions[,c(1:4)]
pancakes <- distinct(pancakes)
sd.all <- aggregate(allregions[,6:18], list(Species=allregions$Species),sd, na.rm=TRUE)
# add in categorical variables
means.all <- cbind(pancakes, means.all[, -1])
rownames(means.all) <- means.all$Species

means.all.long <- pivot_longer(means.all, c(5:17), 
                               names_to = "region", values_to = "density")
remove(pancakes)
```

sort rows to match tree
``` {r}
means.all.sorted <- means.all[match(df.test$Species,means.all$Species),]
# meansnorm.sorted <- meansnorm.sorted[-7,] # removes desmodus from mean dataframe
meansort.all.long <- pivot_longer(means.all.sorted, c(5:(ncol(means.all.sorted))), 
                                  names_to = "region", values_to = "density")

comp.raw <- comparative.data(data = means.all.sorted, phy = batTree, 
                             names.col = "Species", na.omit = FALSE, warn.dropped = TRUE)


```

# region by region models 

proxdist and crancaud
``` {r}
proxdist <- comp.raw %>% pgls(prox ~ dist, data = .) 
summary(proxdist)
plot.pgls(proxdist)

crancaud <- comp.raw %>% pgls(cran ~ caud, data = .)
summary(crancaud)
plot.pgls(crancaud)


```

allpairwise 

```{r}
# 9 regions
pairs <- t(combn(c(1:9), 2, simplify = TRUE)) # makes object with all numerical pairwise comparisons 
pairs[1,1]
pairs[1,2]

# region list to refer to 
regionname <- {c("Elbow_cm2", "Ankleknee.1", "Propat_cm2", "CD_Plagio_cm2", "MD_Plagio_cm2", 
  "CD_Dactylo_Major_cm2", "MD_Dactylo_Major_cm2", "CD_Dactylo_Medius_cm2", "MD_Dactylo_Medius_cm2")}

#  all pairwise comparisons
pairwiseresult <- as.data.frame(t(Reduce(cbind,lapply(1:dim(pairs)[1], function(i) {
    idx1 <- pairs[i,1]
    idx2 <- pairs[i,2]
    name <- paste(regionname[idx1]," vs ", regionname[idx2],sep = "")
    res <- summary(pgls(get(regionname[idx1]) ~ get(regionname[idx2]), data = comp.raw))
    return( c(name, 
              res$coefficients[2,1], 
              res$coefficients[2,4], 
              res$fstatistic[1], 
              res$adj.r.squared) )

} ))))
names(pairwiseresult) <- c("pair", "estimate", "pvalue", "fstatistic", "adjR2")


pairwiseresult %>% subset(pvalue < 0.05/36) %>% write.csv("sigresults.csv") # 36 is number of tests


```


# make nice phylogenetic tree for figure 
```{r}
mar=c(0.1, 1, 0.1, 0.1)
plotTree(batTree,fsize=0.8, ftype="i", lwd=.9)
cols<-setNames(palette()[1:length(unique(means.scaled.sorted$Guild))],sort(unique(means.scaled.sorted$Guild)))
tiplabels(pie = to.matrix(means.scaled.sorted$Guild, 
                          sort(unique(means.scaled.sorted$Guild))), piecol=cols, cex=0.7)
# add.simmap.legend(colors=cols, prompt=TRUE, fsize=0.8)

```

